---
title: "tooth_growth"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lattice)
library(dplyr)

```

# Title

In the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

# Data summary

```{r tooth_growth_data}
library(datasets)
data(ToothGrowth)
summary(ToothGrowth)
str(ToothGrowth)
```

# Data plot

We can also use bwplot to show the tooth growth by factors: supp and dose.

```{r}
boxplot(len~supp*dose, data=ToothGrowth, notch=FALSE, 
  col=(c("gold","darkgreen")),
  main="Tooth Growth", xlab="Suppliment and Dose")
aggregate(ToothGrowth$len,list(ToothGrowth$dose,ToothGrowth$supp)
          ,FUN=function(x) c(x_mean = mean(x), x_sd = sd(x)))
```

## Observation

From the plot above, we can see that when dose=05 or dose=1, the teeth grow differently by using supplment OJ0.5 and VC0.5. We will verify that hypotheis in next section.

# Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

We will only use dose=1 as example to test our hypothesis that tooth growth different by using different supplements.

```{r}
oj1   <- ToothGrowth[which(ToothGrowth$dose==1.0 & ToothGrowth$supp=="OJ" ),1]
vc1   <- ToothGrowth[which(ToothGrowth$dose==1.0 & ToothGrowth$supp=="VC" ),1]
t1_oj_vc <- t.test(oj1, vc1, paired=FALSE, var.equal=TRUE)
t1_oj_vc
```

# Result

We can see a very small p-value < 0.05, then we should reject the hyphothesis and accept the alternative, that different supplements has different effect on the tooth growht when we specifiy the dose=1.0



